Red Wine Exploration by Xiaoyu Chen
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(ggplot2)
library(grid)
library(gridExtra)
library(GGally)
library(psych)
library(dplyr)
library(polycor)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
getwd()
df <- read.csv("wineQualityReds.csv")
```

In this exploration, I will investigate Red Wine data chosen from Data Set Options provided by Udacity. The dataset has 1599 observations with a output variable "quality", which is scored from 3 (worst) to 8 (best), and 11 input variables: fixed.acidity (tartaric acid - g/dm^3), volatile.acidity (acetic acid - g/dm^3), citric.acid (g/dm^3), residual.sugar (g/dm^3), chlorides (sodium chlorides - g/dm^3), free.sulfur.dioxide (mg/dm^3), total.sulfur.dioxide (mg/dm^3), density (g/dm^3), pH, sulphates (potassium sulphate - g/dm^3), and alcohol (% by volume).

Most wines have quality of 5 or 6, fixed.acidity from 6.5 g/dm^3 to 9.5 g/dm^3, volatile.acidity from 0.3 g/dm^3 to 0.8 g/dm^3, citric.acid from 0 g/dm^3 to 0.5 g/dm^3, residual.sugar from 1.5 g/dm^3 to 3 g/dm^3, chlorides between 0.05 g/dm^3 and 0.1 g/dm^3, free.sulfur.dioxide between 3 mg/dm^3 to 15 mg/dm^3, total.sulfur.dioxide from 10 mg/dm^3 to 70 mg/dm^3, density from 0.994 g/dm^3 to 1 g/dm^3, Ph scale between 3 - 4, sulphates from 0.5 g/dm^3 to 0.8 g/dm^3, alcohol from 9% to 11.5%.

# Univariate Plots Section

```{r echo=FALSE, Univariate_Plots1}

# A structure preview of all features in dataset
str(df)

```

```{r echo=FALSE, Univariate_Plots2}

# check if there is missing values acorss all variables in dataset
df[!complete.cases(df),]

```

There is no missing values.

```{r echo=FALSE, Univariate_Plots3}

#delete duplicates
df = distinct(subset(df, select = -c(X)))

#make quality as data type of factor
df$quality = factor(df$quality)

str(df)

```

After removing duplicates, there are 1359 obervatios and 13 variables in the dataset.

All data types look fine right now. Let's take a look at summary.

```{r echo=FALSE, Univariate_Plots4}

#A summary preview of all features
summary(df)

```

The quality range is from 3 (worst) to 8 (best), and most wines have quality of 5 or 6. let's explore each variables one by one.

Next, we will examine variable distributions one by one.

```{r echo=FALSE, Univariate_Plots5}

#"quality" (output variable) distribution
ggplot(aes(quality), data = df) +
  geom_bar()

```

The most wines have quality of 5 or 6. The best red wines and the worst wines have least counts compared with other quality scores.

```{r echo=FALSE, Univariate_Plots6}

#fixed.acidity distribution
p1 <- ggplot(aes(fixed.acidity), data = df) +
  geom_histogram(color = 'black', binwidth = 0.1) +
  scale_x_continuous(breaks = seq(4, 16, 0.5))

# scale log10 transformation of fixed.acidity
p2 <- ggplot(aes(fixed.acidity), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01) +
  scale_x_log10(breaks = seq(4, 16, 0.5))

grid.arrange(p1, p2)

```

The "fixed.acidity" is transformed to log distribution because of the long right tail. The transformed distribution peaks around 7.2. Most wines have fixed.acidity from 6.5 to 9.5.

```{r echo=FALSE, Univariate_Plots7}

#volatile.acidity distribution
p1 <- ggplot(aes(volatile.acidity), data = df) +
  geom_histogram(color = 'black', binwidth = 0.02)

p2 <- ggplot(aes(volatile.acidity), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01) +
  scale_x_log10(breaks = seq(0,1.6, 0.1))

grid.arrange(p1, p2)

```

The distribution of  transformed volatile.acidity peaks about 0.6. Most of volatile.acidity are from 0.3 to 0.8. Since high levels of volatile.acidity can lead to unpleasant taste, I wonder what does the quality distribution look like when given high volatile.acidity over 0.8.

```{r echo=FALSE, Univariate_Plots8}

#citric.acid distribution
ggplot(aes(citric.acid), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01)

```

The distribution of citric.acid looks linearly decreases and peaks at 0. Since the citric.acid can add freshness and flavor to wines, I wonder what are the qualities of these wines with 0 citric.acid.

```{r echo=FALSE, Univariate_Plots9}

p1 <- ggplot(aes(residual.sugar), data = df) +
  geom_histogram(color = 'black', binwidth = 0.1) +
  scale_x_continuous(breaks = seq(0,16,1))

p2 <- ggplot(aes(residual.sugar), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01) +
  scale_x_log10(breaks = seq(0,16,1))

grid.arrange(p1, p2)

```

The historgram of residual.sugar is right skewed so it is log transfromed. The transformed residual.sigar peaks at 2, and most wines in the range between 1.5 and 3. 

Because it's very rare to find wines with less than 1g/liter residual.sugar, we take a look of them in the next.

```{r echo=FALSE, Univariate_Plots10}

#subset residual.sugar < 1
subset(df, residual.sugar < 1)

```

It looks all numbers are fine in this row. So we don't consider it as outlier.

```{r echo=FALSE, Univariate_Plots11}

#chlorides distribution
p1 <- ggplot(aes(chlorides), data = df) +
  geom_histogram(color = 'black', binwidth = 0.005) +
  scale_x_continuous(breaks = seq(0,0.7,0.05))

#log transformed
p2 <- ggplot(aes(chlorides), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01) +
  scale_x_log10(breaks = seq(0,0.7,0.05))

grid.arrange(p1, p2)

```

The transformed "chlorides" peaks around 0.8, and most wines have chlorides between 0.05 and 0.1.

```{r echo=FALSE, Univariate_Plots12}

#free.sulfur.dioxide distribution and log transformation
p1 <- ggplot(aes(free.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', binwidth =0.01) +
  scale_x_continuous(breaks = seq(0,80,5))

p2 <- ggplot(aes(free.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', binwidth = 0.01) +
  scale_x_log10(breaks = seq(0, 80,5))

grid.arrange(p1, p2)

```

Most free.sulfur.dioxide are intergers. After transformed,free.sulfur.dioxide peaks at 6 and most wines have free.sulfur.dioxide between 3 to 15.

We know that at free SO2 concentrations over 50 ppm, SO2 becomes evident in the nose and taste of wine. So we create a new variable "Free_SO2_50" to represent whether free SO2 concentration greater than 50 ppm or not. The free SO2 concentration is calculated by using (mass of free SO2) / mass, in other words, (1L times free.sulfur.dioxide) / ( 1L times density).

```{r echo=FALSE, Univariate_Plots13}

#create new variable : Free_SO2_50, if free SO2 > 50, then the value = 1, if not, the value = 0
df$Free_SO2_50 <- ifelse((df$free.sulfur.dioxide / df$density) >= 50, 'free.sulfur.dioxide>50ppm','free.sulfur.dioxide < 50ppm')
df$Free_SO2_50 <- factor(df$Free_SO2_50)
ggplot(aes(Free_SO2_50), data = df) +
  geom_bar()

```

Most wines have free SO2 less than 50 ppm, but there are some over 50 ppm. I wonder what are the quality scores of wines having free SO2 over 50 ppm.

```{r echo=FALSE, Univariate_Plots14}

#total.sulfur.dioxide distribution and log transformed
p1 <- ggplot(aes(total.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', binwidth = 0.2) +
  scale_x_continuous(breaks = seq(0, 300, 20))

p2 <- ggplot(aes(total.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', , binwidth = 0.005) +
  scale_x_log10(breaks = seq(0, 300, 10))

grid.arrange(p1, p2)


```

Most total.sulfur.dioxide are intergers and in the range from 10 to 70. Since total.sulfur.dioxide is sum of free and bound SO2, so we can make a new variable bound.sulfur.dioxide (mg/dm^3) to investigate its effects about red wine quality.

```{r echo=FALSE, Univariate_Plots15}

#create new variable : bound.sulfur.dioxide
df$bound.sulfur.dioxide <- df$total.sulfur.dioxide - df$free.sulfur.dioxide

```

Let's take a look at the distribution of the new created variable "bound.sulfur.dioxide"

```{r echo=FALSE, Univariate_Plots16}

p1 <- ggplot(aes(bound.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', binwidth = 1) +
  scale_x_continuous(breaks = seq(0, 300, 20))

p2 <- ggplot(aes(bound.sulfur.dioxide), data = df) +
  geom_histogram(color = 'black', binwidth = 0.005) +
  scale_x_log10(breaks = seq(0, 300, 10))

grid.arrange(p1, p2)

```

The bound.sulfur.dioxide peaks around 9. Compared with having free.sulfur.dioxide, wines contain more bound.sulfur.dioxide. I guess bound.sulfur.dioxide has very high correlation with free.sulfur.dioxide and total.sulfur.dioxide.

```{r echo=FALSE, Univariate_Plots17}

#density distribution
ggplot(aes(density), data = df) +
  geom_histogram(color = 'black', binwidth =  0.0001) +
  scale_x_continuous(breaks = seq(0.99, 1.1, 0.001))

```

The distribution of density is normal and most of density is ranging from 0.994 to 1. The density is related to sugar and alcohol, we will check their relations in the bivariate analysis.

```{r echo=FALSE, Univariate_Plots18}

#pH distribution
ggplot(aes(pH), data = df) +
  geom_histogram(color = 'black', binwidth = 0.005) +
  scale_x_continuous(breaks = seq(2.5, 4.1, 0.1))

```

The PH distribution is normal. Most wines are between 3 - 4 on the Ph scale, sp I want to check rows with pH less than 3 in the next.

```{r echo=FALSE, Univariate_Plots19}

#subset pH < 3
subset(df, pH < 3)

```

The all rows have very high values of sulfur.dioxide or citric.acid or fixed.acidity. It looks that pH is related to sulfur.dioxide and acidity.

```{r echo=FALSE, Univariate_Plots20}

#sulphates distribution
p1 <- ggplot(aes(sulphates), data = df) +
  geom_histogram(color = 'black', binwidth = 0.005) +
  scale_x_continuous(breaks = seq(0, 2, 0.1))
#log transformation
p2 <- ggplot(aes(sulphates), data = df) +
  geom_histogram(color = 'black', binwidth = 0.005) +
  scale_x_log10(breaks = seq(0, 2, 0.1))

grid.arrange(p1, p2)

```

Most wines have sulphates between 0.5 to 0.8. I guess sulphates have close correlation with free.sulfur.dioxide.

```{r echo=FALSE, Univariate_Plots21}

#alcohol distribution
p1 <- ggplot(aes(alcohol), data = df) + 
  geom_histogram(color = 'black', binwidth =  0.05) +
  scale_x_continuous(breaks = seq(0,15, 0.5))

p2 <- ggplot(aes(alcohol), data = df) + 
  geom_histogram(color = 'black', binwidth =  0.001) +
  scale_x_log10(breaks = seq(0,15, 0.5))

grid.arrange(p1, p2)

```

Alcohol peaks around 9.5 and the count of wines are decreasing as alcohol increasing more than 9.5. Most wines have alcohol ranging from 9% to 11.5%.

After displaying every single variable, we found that all attributes in red wines could be concluded in:acidity, sulfur dioxide, sugar, salt, and alcohol. I guess that density has close clorrelation with sugar, salt and alcohol. The pH value has close correlation with acidity and probably the concentration of sulfur dioxide. 

# Univariate Analysis

### What is the structure of your dataset?

There are 1359 unique wines in the dataset with 11 attributes (fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol). All of attributes are numeric data.

### What is/are the main feature(s) of interest in your dataset?

The main features in the dataset are alcohol and quality. 

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

I guess fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, and sulphates will help support my investigation.

### Did you create any new variables from existing variables in the dataset?

I create two new variables in total.

The first is called "Free_SO2_50", which is to represent whether free SO2 concentration greater than 50 ppm or not. If "Free_SO2_50" equals to or is greater than 50, we get "1", otherwise we get "0". That is a factor variable. 

The second is "bound.sulfur.dioxide" (mg/dm^3), which is created by using total.sulfur.dioxide minus free.sulfur.dioxide. 

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I remove duplicated rows before display structure of dataset. Because I think same wines could not help us more to investigate the dataset.

I use log10 to transform almost all features beacuase of right skewed. 

It's very rare to find wines with less than 1g/liter residual.sugar, but we do have a wine with 0.9. I list the row out to see if it is a outlier. Since other variables look fine, so I keep it.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots1, warning=FALSE}
#correlation matrix
hetcor(df)

```

Residual.sugar only moderately correlates with density, and chlorides only correlates with sulphates. There is no other obvious correlations in both residual.sugar and chlorides.

One interestin finding is that, all sulfur dioxide (free, bound and total) only highly correlated with each other, and have no meaningful corelations with other variables, including quality. Both bound and free sulfur dioxide have large correlations with total sulfur dioxide, which makes sense. Bound sulfur dioxide and free are moderately correlated.

Another interesting finding is that although "sulphates" can contribute to sulfur dioxide gas levels, the correlation between them is much weaker than I expected. 

Because of the previous conclusions, let's subset data with variables excluding "residual.sugar", "chlorides", "free.sulfur.dioxide", "total.sulfur.dioxide", "bound.sulfur.dioxide" and "Free_SO2_50".

```{r echo=FALSE, Bivariate_Plots2}

pairs.panels(subset(df, select = -c(residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, bound.sulfur.dioxide, Free_SO2_50 )))

```

Quality has moderate correlations with alcohol and Volatile.acidity, and slight correlation with Sulphates and citric.acidity.

One unexpected finding is that density is highly correlated with fixed.acidity. At the same time, density has no obvious correlation with volatile.acidity and meaningful correlation with citric.acid.

We will look closer at plots about quality and some other variables, including alcohol, volatile.acidity, sulphates, citric acidity, density, bound.sulfur.dioxide and Free_SO2_50. 

```{r echo=FALSE, Bivariate_Plots3}

#scatter plot  and box plot between quality and alcohol
p1 <- ggplot(df, aes(alcohol, quality)) +
  geom_jitter(alpha = 0.1)

p2 <- ggplot(aes(quality, alcohol),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)

```

```{r echo=FALSE, Bivariate_Plots4}

#statistics of alcohol in each quality score 

by(df$alcohol, df$quality, summary)

```

As alcohol increases, the quality gets more chance at higher levels. Quality of 5 looks have lowest median of alcohol, and it also has least variance in the middle 50%.

We also notice that many wines have same levels of alcohol at different quality. 

```{r echo=FALSE, Bivariate_Plots5}

#linear regression between quality and alcohol

df$quality <- as.numeric(levels(df$quality))[df$quality]

fit <- lm(quality ~ alcohol, data = df)
summary(fit)

```

Based on the R^2 value, alcohol can explain 23% of the variance in quality.

```{r echo=FALSE, Bivariate_Plots6}

#scatter plot  and box plot between quality and volatile.acidity
df$quality =factor(df$quality)

p1 <- ggplot(df, aes(volatile.acidity, quality)) +
  geom_jitter(alpha = 0.1)

p2 <- ggplot(aes(quality, volatile.acidity),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)


```

```{r echo=FALSE, Bivariate_Plots7}

##statistics of volatile.acidity in each quality score

by(df$volatile.acidity, df$quality, summary)

```

As volatile.acidity increases, the quality gets less chance at higher levels. The means of alcohol show a decreasing trend from quality score of 3 to 8.
Similarly, We also notice that many wines having same levels of volatile.acidity are scored differently. 

```{r echo=FALSE, Bivariate_Plots8}

#linear regression between quality and volatile.acidity

df$quality <- as.numeric(levels(df$quality))[df$quality]

fit <- lm(quality ~ volatile.acidity, data = df)
summary(fit)

```

It looks that volatile.acidity has a little worse explaination power than alcohol on quality. There is 16% of the variance in quality can be explained by volatile.acidity.

```{r echo=FALSE, Bivariate_Plots9}

#scatter plot  and box plot between quality and sulphates
df$quality <- factor(df$quality)

p1 <- ggplot(df, aes(sulphates, quality)) +
  geom_jitter(alpha = 0.1)


p2 <- ggplot(aes(quality, sulphates),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)

```

```{r echo=FALSE, Bivariate_Plots10}

#scatter plot  and box plot between quality and alcohol

by(df$sulphates, df$quality, summary)

```

As sulphates increase, quality tends to be higher. Quality of 7 and 8 have same median of sulphates, but 8 has less variance than 7.

Many wines having same levels of sulphates are scored differently.

```{r echo=FALSE, Bivariate_Plots11}

#linear regression between quality and sulphates

df$quality <- as.numeric(levels(df$quality))[df$quality]

fit <- lm(quality ~ sulphates, data = df)
summary(fit)

```

 There is 6% of the variance in quality can be explained by sulphates.

```{r echo=FALSE, Bivariate_Plots12}

#scatter plot  and box plot between quality and citric.acid
df$quality <- factor(df$quality)

p1 <- ggplot(df, aes(citric.acid, quality)) +
  geom_jitter(alpha = 0.1)

p2 <- ggplot(aes(quality, citric.acid),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)

```

```{r echo=FALSE, Bivariate_Plots13}

#scatter plot  and box plot between quality and alcohol

by(df$citric.acid, df$quality, summary)

```

Quality and citric.acid have slight correlation from scatter plot. The means and medians of citric.acid are having upward trends as quality increases.

We also found that all wines with 0 citric.acid could be any scores except 8. 

```{r echo=FALSE, Bivariate_Plots14}


#scatter plot  and box plot between quality and density
p1 <- ggplot(df, aes(density, quality)) +
  geom_jitter(alpha = 0.1)

p2 <- ggplot(aes(quality, density),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)

```

```{r echo=FALSE, Bivariate_Plots15}

#scatter plot  and box plot between quality and alcohol

by(df$density, df$quality, summary)

```

Quality tends to decrease as density increases.The means and medians of density are decreasing as quality increases overall, even though the difference is slight. 

```{r echo=FALSE, Bivariate_Plots16}

#scatter plot  and box plot between quality and bound.sulfur.dioxide
p1 <- ggplot(df, aes(bound.sulfur.dioxide, quality)) +
  geom_jitter(alpha = 0.1)

p2 <- ggplot(aes(quality, bound.sulfur.dioxide),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point')

grid.arrange(p1, p2)


```

```{r echo=FALSE, Bivariate_Plots17}

#scatter plot  and box plot between quality and alcohol

by(df$bound.sulfur.dioxide, df$quality, summary)

```

It looks we only see slight correlation between qualtity and bound.sulfur.dioxide. The meams and medians are increasing then decreasing as quality increases.

```{r echo=FALSE, Bivariate_Plots18}

#quality scores for Free_SO2_50
ggplot(df, aes(quality)) +
  geom_bar() +
  facet_wrap(~Free_SO2_50)

```

The wines with more than 50 ppm of free.sulfur.dioxide are having not bad quality scores from 5 to 7.

Next, I'll check how the alcohol varies with some variables that may have meaningful correlations with alcohol. These variables are: volatile.acidity, chlorides, density and PH.

```{r echo=FALSE, Bivariate_Plots19}

#scatter plot between alcohol and volatile.acidity
ggplot(df, aes(volatile.acidity, alcohol)) +
  geom_point(alpha = 0.1)

```

Although both alcohol and volatile.acidity have moderate and almost same correlations with qualtiy, the scatter plot shows the lack of correlations between alcohol and volatile.acidity.

```{r echo=FALSE, Bivariate_Plots20}

#scatter plot between alcohol and chlorides
ggplot(df, aes(chlorides, alcohol)) +
  geom_point(alpha = 0.1) 

```

Let's take a close look at chlorides between 0 to 0.2.

```{r echo=FALSE, Bivariate_Plots21, , warning=FALSE}

ggplot(df, aes(chlorides, alcohol)) +
  geom_point(alpha = 0.1) +
  scale_x_continuous(limits = c(0,0.2))

```

There is slight correlations between chlorides and alcohol. Alcohol is rarely greater than 12 when chlorides are more than 0.125.

```{r echo=FALSE, Bivariate_Plots22}

#scatter plot between alcohol and density
ggplot(df, aes(density, alcohol)) +
  geom_point(alpha = 0.1)

```

It looks clearly that density and alcohol are negatively correlated.

```{r echo=FALSE, Bivariate_Plots23}

#scatter plot between alcohol and pH
ggplot(df, aes(pH, alcohol)) +
  geom_point(alpha = 0.1)



```

It shows a slight positive correlations between alcohol and pH. 

Next, let's look at the correlation among acidity:fixed.acidity, volatile.acidity, and citric.acid. And I will also check their correlations with pH.

```{r echo=FALSE, Bivariate_Plots24}

#the correlation among acidity:fixed.acidity, volatile.acidity, and citric.acid
p1 <- ggplot(df, aes(fixed.acidity, volatile.acidity)) +
  geom_point(alpha = 0.1)

p2 <- ggplot(df, aes(fixed.acidity, citric.acid)) +
  geom_point(alpha = 0.1)

p3 <- ggplot(df, aes(citric.acid, volatile.acidity)) +
  geom_point(alpha = 0.1)

grid.arrange(p1, p2, p3)

```

Citic.acid is strongly correlated with both volatile.acidity and fixed.acidity. The relationship between fixed.acidity and citric.acid is tend to be exponential rather than linear. There is slight correlation between volatile and fixed.

Next, let's look at their correlations with pH.

```{r echo=FALSE, Bivariate_Plots25}

#the correlation between acidity and pH
p1 <- ggplot(df, aes(fixed.acidity, pH)) +
  geom_point(alpha = 0.1)

p2 <- ggplot(df, aes(volatile.acidity, pH)) +
  geom_point(alpha = 0.1)

p3 <- ggplot(df, aes(citric.acid, pH)) +
  geom_point(alpha = 0.1)

grid.arrange(p1, p2, p3)


```

Both citic.acid and fixed.acidity are negatively correlated with pH which makes sense. At the same time, volatile.acidity looks positively correlated with pH, which is also reasonable.

Next, let's explore density with relations to fixed.acidity, sugar, chlorides and pH.

```{r echo=FALSE, Bivariate_Plots26}

#the correlation between density and fixed.acidity
ggplot(df, aes(fixed.acidity, density)) +
  geom_point(alpha = 0.1)

```

```{r echo=FALSE, Bivariate_Plots27}

#linear regression between fixed.acidity and density

fit <- lm(density ~ fixed.acidity, data = df)
summary(fit)

```

Density is highly correlated with fixed.acidity. About 45% of density could be explained by fixed.acidity based on R^2.

```{r echo=FALSE, Bivariate_Plots28}

#the correlation between density and residual.sugar
ggplot(df, aes(residual.sugar, density)) +
  geom_point(alpha = 0.1)

```

Let's zoom in the scatter plot.

```{r echo=FALSE, Bivariate_Plots29, warning=FALSE}

ggplot(df, aes(residual.sugar, density)) +
  geom_jitter(alpha = 0.1) +
  scale_x_continuous(limits = c(1,4))

```

Density goes upward as residual.sugar increases and that makes sense. We also notice there are vertical bands along residual.sugar values.

```{r echo=FALSE, Bivariate_Plots30, warning=FALSE}

#the correlation between density and chlorides
ggplot(df, aes(chlorides, density)) +
  geom_point(alpha = 0.1) + 
  scale_x_continuous(limits = c(0,0.2))

```

Similarly, density increases as chlorides increase. That makes sense since sugar and salt could increase density in normal life. 

```{r echo=FALSE, Bivariate_Plots31}

#the correlation between density and pH
ggplot(df, aes(pH, density)) +
  geom_point(alpha = 0.1)


```

Density tends to be lower as pH grows. 

Next we will check sulphates with free.sulfur.dioxide, volatile.acidity and chlorides.

```{r echo=FALSE, Bivariate_Plots32}

#scatter plot between sulphates and free.sulfur.dioxide
ggplot(df, aes(free.sulfur.dioxide, sulphates)) +
  geom_jitter(alpha = 0.1)

```

Like mentioned before, sulphates and free.sulfur.dioxide are still lack of correlations from scatter plot.

```{r echo=FALSE, Bivariate_Plots33}

#scatter plot between sulphates and volatile.acidity
ggplot(df, aes(volatile.acidity, sulphates)) +
  geom_jitter(alpha = 0.1)

```

There is a slight correlation between volatile.acidity and sulphates.

```{r echo=FALSE, Bivariate_Plots34}

#scatter plot between sulphates and volatile.acidity
ggplot(df, aes(chlorides, sulphates)) +
  geom_jitter(alpha = 0.1) 

```

Let's take a closer look.

```{r echo=FALSE, Bivariate_Plots35, warning=FALSE}

ggplot(df, aes(chlorides, sulphates)) +
  geom_jitter(alpha = 0.1) +
  scale_x_continuous(limits = c(0,0.2))

```

Sulphates and chlorides have a correlation 0.3946. Overall, sulphates grows as chlorides increases. However, if we take a closer look at the most data in the middle, it looks hard to find relation between them.


# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

Alcohol and volatile.acidity moderately correlate with quality. As alcohol increases, the quality gets more chance to be higher from scatter plot. On the contrary, the quality gets less chance at higher levels when volatile.acidity increases. We also notice that many wines having same levels of volatile.acidity or alcohol are scored differently.

Based on R^2 values, alcohol can explain 23% of the variance in quality and volatile.acidity explains 16%.

There is no any obvious correlation between alcohol and volatile.acidity from scatter plot.

In addition, there is a correlation between sulphates and quality as well. Citric.acid and density only show slight correlation with quality.

One intersting finding is the wines with more than 50 ppm of free.sulfur.dioxide are having not bad quality scores from 5 to 7.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Both bound and free sulfur dioxide have large correlations with total sulfur dioxide, which makes sense. Bound sulfur dioxide and free are moderately correlated.

Alcohol nagatively correlates to density, slightly positively correlates to pH, and slightly negatively to chlorides.

Citric.acid correlates with both fixed.acidity and volatile.acidity. The relationship between citric.acid and fixed.acidity is exponential rather than linear.

Both citric.acid and fixed.acidity negatively correlate with pH, and volatile.acidity positively correlates with pH.

Density correlates with fixed.acidity the most. At the same time, density clearly correlates with chlorides, slightly correlates with residual.sugar and pH.

Sulphates don't show correlation with free.sulfur.dioxide, but the scatter plot displays slight negative correlation between sulphates and volatile.acidity.

### What was the strongest relationship you found?

Fixed.acidity is strongly and positively coreelated with density. About 45% of density could be explained by fixed.acidity based on R^2.

# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots1}

ggplot(data = df, aes(x = alcohol, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_density()
  

```

```{r echo=FALSE, Multivariate_Plots2}

ggplot(data = df, aes(x = volatile.acidity, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_density()

```

```{r echo=FALSE, Multivariate_Plots3}

ggplot(data = df, aes(x = sulphates, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_density()

```

```{r echo=FALSE, Multivariate_Plots4}

ggplot(data = df, aes(x = citric.acid, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_density()

```

```{r echo=FALSE, Multivariate_Plots5}

ggplot(data = df, aes(x = density, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_density()

```

Wines with higher levels of alcohol, sulphates and citric.acid, and lower levels of volatile.acidity and density tend to occur more often at higher quality scores.

```{r echo=FALSE, Multivariate_Plots6}

ggplot(df, aes(x = alcohol, y = volatile.acidity, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_point()

```

```{r echo=FALSE, Multivariate_Plots7}

ggplot(df, aes(x = alcohol, y = volatile.acidity)) +
  geom_point(alpha = 0.2) + 
  facet_wrap(~quality)

```

Quality scores of 4, 5, 6, 7 cluster by volatile.acidity and alcohol values, but quality scores of 3 and 8 don't show much cluster trend from the plot. This is probably because there is very few amount of data in the 3 and 8.

We also notice that better qualities (6, 7, 8) have very few wines with volatile.acidity more than 0.8. In addition, quality scores of 3, 4, 5 have very few wines with alcohol greater than 12.

```{r echo=FALSE, Multivariate_Plots8}

ggplot(df, aes(x = alcohol, y = sulphates, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_point()

```

This plot shows that if we keep alcohol constant, the wines with more sulphates tend to have higher quality scores.

```{r echo=FALSE, Multivariate_Plots9}

ggplot(df, aes(x = sulphates , y = chlorides, color = quality)) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  geom_point()

```

This plot shows that if we keep sulphates constant, the wines with less chlorides tend to have lower quality scores.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

Wines with higher levels of alcohol, sulphates and citric.acid, and lower levels of volatile.acidity and density tend to occur more often at higher quality scores.

Better qualities (6, 7, 8) have very few wines with volatile.acidity more than 0.8. In addition, quality scores of 3, 4, 5 have very few wines with alcohol greater than 12.

If we keep alcohol constant, the wines with more sulphates tend to have higher quality scores.

If we keep sulphates constant, the wines with less chlorides tend to have lower quality scores.

### Were there any interesting or surprising interactions between features?

Quality scores of 5, 6, 7 cluster by volatile.acidity and alcohol values, and quality scores of 3, 4, and 8 not cluster. It is probably because that there are different amount of wines in each quality scores. 

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

#"quality" (output variable) distribution
ggplot(aes(quality), data = df) +
  geom_bar() +
  ylab ('Number of Wines') +
  ggtitle('quality Distribution')

```

### Description One

The most wines have quality of 5 or 6. The best red wines and the worst wines have least counts compared with other quality scores.

### Plot Two
```{r echo=FALSE, Plot_Two}

p1 <- ggplot(aes(quality, alcohol),data = df) +
  geom_boxplot() +
  stat_summary(fun.y =mean,col='red',geom='point') +
  ylab ('alcohol (% by volume)') +
  ggtitle ('Wine quality by alcohol')

p2 <- ggplot(data = df, aes(x = alcohol, color = quality)) +
  geom_density()+
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality levels')) +
  scale_x_continuous(breaks = seq(8,15,1)) +
  xlab ('alcohol (% by volume)') +
  ggtitle ('Density of wine quality by alcohol')

grid.arrange(p1, p2)


```

### Description Two

The means and medians of alcohol have increasing trends overall as quality increases. Wines having more alcohol tend to occur more often at higher quality scores. The wines with quality 5 have lowest alcohol median and mean, and they have least alcohol variance in the middle 50%.

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(df, aes(x = alcohol, y = volatile.acidity)) +
  geom_point(alpha = 0.2) + 
  facet_wrap(~quality) +
  xlab ('alcohol (% by volume)') +
  ylab ('volatile.acidity (g / dm^3)')+
  ggtitle ('Facet wrap by quality')
  

```

### Description Three

Quality scores of 4, 5, 6, 7 cluster by volatile.acidity and alcohol values, but quality scores of 3 and 8 don't show much cluster trend from the plot.

We also notice that better qualities (6, 7, 8) have very few wines with volatile.acidity more than 0.8. In addition, quality scores of 3, 4, 5 have very few wines with alcohol greater than 12.

------

# Reflection

Summary: This exploration is to investigate the relationship between quality scores for wines and their attributes, like acidity, sulfur dioxide, sugar, salt, and alcohol. We found that alcohol has a positive correlation with quality and volatile.acidity as a negative correlation with quality. Although both alcohol and volatile.acidity has almost same correlation coefficient if we ignore positive and negative sign, it looks that volatile.acidity has a little worse explaination power than alcohol on quality. Better qualities (6, 7, 8) have very few wines with volatile.acidity more than 0.8. In addition, quality scores of 3, 4, 5 have very few wines with alcohol greater than 12. There is a lack of correlation between alcohol and volatile.acidity.

Furthermore, Wines with higher levels of alcohol, sulphates and citric.acid, and lower levels of volatile.acidity and density tend to occur more often at higher quality scores.

One interesting finding is that although free SO2 concentrations over 50 ppm may results that SO2 becomes evident in the nose and taste of wine, they still get not bad quality from 5 to 7.

Another interesting finding is that wines with 0 citric.acid could be in any quality scores except 8. 

The struggle is it is the first time for me to handle these many variables at the same time. Sometimes it is a little hard to organize them properly. Until in the middle of analysis, I found all variables could be concluded in: acidity, sulfur dioxide, sugar, salt, and alcohol. Then it is much more clear for me to analyze them.

One limitation about this analysis is we didn't build any model for quality. Alcohol and volatile.acidity should be the most important components in the model for quality. For the future work in building model, I would try other variables like sulphate or citricacid to improve precision of model.

# Reference

https://www.udacity.com

https://stackoverflow.com

